---
title: "Second MedBench paper: PDQI+ campaign analysis"
author: "Max Gordon"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    dev: pdf
    self_contained: false
params:
  campaign_file: "data/output/reviewCampaign_<slug>.json"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.dpi = 300)
library(tidyverse)
library(jsonlite)
library(glue)
library(boot)

pdqi_domains <- c(
  "accurate",
  "thorough",
  "useful",
  "clear",
  "organized",
  "succinct",
  "synthesizes",
  "internallyConsistent",
  "upToDate"
)

pdqi_labels <- c(
  accurate = "Accurate",
  thorough = "Thorough",
  useful = "Useful / relevant",
  clear = "Clear / understandable",
  organized = "Organized",
  succinct = "Succinct",
  synthesizes = "Synthesizes / prioritizes",
  internallyConsistent = "Internally consistent",
  upToDate = "Up-to-date"
)
```

```{r helpers}
convert_decimal_to_lancet <- function(x) {
  x |>
    map_chr(\(value) str_replace(value, "\\.", "&middot;"))
}

txt_round <- function(x, digits = 1) {
  ifelse(is.na(x), "", htmlTable::txtRound(x, digits = digits) |> convert_decimal_to_lancet())
}

style_table <- function(x) {
  if (requireNamespace("kableExtra", quietly = TRUE)) {
    return(kableExtra::kable_styling(x, full_width = FALSE, bootstrap_options = c("striped", "hover")))
  }
  x
}

mean_ci <- function(x, digits = 1) {
  x <- x[!is.na(x)]
  if (length(x) == 0) {
    return("")
  }
  if (length(x) < 2) {
    return(txt_round(mean(x), digits = digits))
  }
  boot_stat <- function(data, indices) mean(data[indices], na.rm = TRUE)
  boot_result <- boot(x, boot_stat, R = 1000)
  ci <- tryCatch(boot.ci(boot_result, type = "perc")$percent[c(4, 5)],
                 error = \(e) c(NA_real_, NA_real_))
  glue("{txt_round(mean(x), digits)} (95% CI {txt_round(ci[1], digits)} to {txt_round(ci[2], digits)})")
}

safe_factor <- function(x, ref = NULL) {
  out <- factor(x)
  if (!is.null(ref) && ref %in% levels(out)) {
    out <- relevel(out, ref = ref)
  }
  out
}

normalise_language <- function(x) {
  case_when(
    x == "original" ~ "English",
    is.na(x) | x == "" ~ "Unknown",
    TRUE ~ x
  )
}

normalise_generator <- function(generator_type, generated_by) {
  case_when(
    generator_type == "HUMAN" | generated_by == "Human" ~ "Human",
    generator_type == "LLM" ~ "LLM",
    str_starts(generated_by, "\\$") ~ "LLM",
    TRUE ~ "Unknown"
  )
}

normalise_prompt <- function(generator, prompt_type, generated_by) {
  case_when(
    generator == "Human" ~ "Human",
    str_detect(generated_by, "@basic") ~ "Single",
    str_detect(generated_by, "@decompos") ~ "Multistep",
    !is.na(prompt_type) & prompt_type != "" ~ prompt_type,
    TRUE ~ "Unknown"
  )
}

output_mixed_model <- function(model, outcome, digits = 2) {
  if (!requireNamespace("broom.mixed", quietly = TRUE)) {
    return(glue("Mixed model for {outcome} was fitted, but broom.mixed is not installed for table output."))
  }

  broom.mixed::tidy(model, conf.int = TRUE) |>
    filter(is.na(group)) |>
    transmute(
      Term = term,
      Estimate = txt_round(estimate, digits),
      `95% CI` = glue("{txt_round(conf.low, digits)} to {txt_round(conf.high, digits)}")
    ) |>
    knitr::kable("html", caption = glue("Mixed model results for {outcome}"), escape = FALSE) |>
    style_table()
}

can_fit_mixed_model <- function(data, outcome) {
  data |>
    filter(!is.na(.data[[outcome]]), !is.na(generator), !is.na(reviewerName)) |>
    summarise(
      rows = n(),
      generators = n_distinct(generator),
      reviewers = n_distinct(reviewerName),
      token_values = n_distinct(chartTokenCount[!is.na(chartTokenCount)]),
      .groups = "drop"
    ) |>
    with(rows >= 8 && generators >= 2 && reviewers >= 2 && token_values >= 2)
}

fit_quality_model <- function(data, outcome) {
  if (!requireNamespace("lme4", quietly = TRUE)) {
    return(NULL)
  }

  model_data <- data |>
    filter(!is.na(.data[[outcome]]), !is.na(generator), !is.na(reviewerName)) |>
    mutate(
      generator = safe_factor(generator, ref = "Human"),
      chartTokenCountScaled = as.numeric(scale(chartTokenCount))
    )

  if (!can_fit_mixed_model(model_data, outcome)) {
    return(NULL)
  }

  lme4::lmer(as.formula(glue("{outcome} ~ generator + chartTokenCountScaled + (1 | reviewerName)")),
             data = model_data)
}
```

```{r load-data}
if (!file.exists(params$campaign_file)) {
  stop(glue("Campaign export not found: {params$campaign_file}. Run download_platform_review_campaign.py first."))
}

campaign_export <- read_file(params$campaign_file) |>
  fromJSON(simplifyVector = TRUE, flatten = TRUE)

campaign <- campaign_export$campaign

reviews <- campaign_export$reviewsFlat |>
  as_tibble() |>
  mutate(
    language = normalise_language(language),
    generator = normalise_generator(summaryGeneratorType, summaryGeneratedBy),
    promptType = normalise_prompt(generator, promptType, summaryGeneratedBy),
    across(c(overall, pdqi9_mean, pdqi9_total, safety, bias, hallucinationSeverity,
             chartTokenCount, chartWordCount, noteCount, medicationCount, labValueCount,
             factCount, lengthOfStayDays, seconds2review),
           as.numeric),
    across(starts_with("pdqi_") & !ends_with("__comment"), as.numeric)
  )

pdqi_cols <- paste0("pdqi_", pdqi_domains)

chart_lengths <- reviews |>
  distinct(chartTranslationId, caseName, specialty, language, chartTokenCount, chartWordCount,
           noteCount, medicationCount, labValueCount, factCount, lengthOfStayDays)
```

# Campaign

This report analyses review campaign `r campaign$label` (`r campaign$id`). The campaign review type is `r campaign$reviewType`.

```{r campaign-summary}
tibble(
  Metric = c("Campaign label", "Review type", "Charts", "Summaries", "Reviews", "Reviewers", "Cases", "Languages", "Specialties"),
  Value = c(
    campaign$label,
    campaign$reviewType,
    length(campaign$charts$id),
    sum(map_int(campaign$charts$summaries, nrow)),
    nrow(reviews),
    n_distinct(reviews$reviewerName),
    n_distinct(reviews$caseName),
    paste(sort(unique(reviews$language)), collapse = ", "),
    paste(sort(unique(reviews$specialty)), collapse = ", ")
  )
) |>
  knitr::kable("html", escape = FALSE, caption = "Campaign overview") |>
  style_table()
```

```{r reviewer-contributions}
reviews |>
  count(reviewerName, sort = TRUE, name = "Reviews") |>
  rename(Reviewer = reviewerName) |>
  knitr::kable("html", caption = "Reviewer contributions") |>
  style_table()
```

# Chart Length

```{r chart-length-table}
chart_lengths |>
  group_by(specialty, language) |>
  summarise(
    Charts = n(),
    `Chart tokens, median (range)` = glue(
      "{median(chartTokenCount, na.rm = TRUE)} ({min(chartTokenCount, na.rm = TRUE)}-{max(chartTokenCount, na.rm = TRUE)})"
    ),
    `Chart words, median (range)` = glue(
      "{median(chartWordCount, na.rm = TRUE)} ({min(chartWordCount, na.rm = TRUE)}-{max(chartWordCount, na.rm = TRUE)})"
    ),
    `Notes, median (range)` = glue(
      "{median(noteCount, na.rm = TRUE)} ({min(noteCount, na.rm = TRUE)}-{max(noteCount, na.rm = TRUE)})"
    ),
    `Length of stay, median days (range)` = glue(
      "{median(lengthOfStayDays, na.rm = TRUE)} ({min(lengthOfStayDays, na.rm = TRUE)}-{max(lengthOfStayDays, na.rm = TRUE)})"
    ),
    .groups = "drop"
  ) |>
  knitr::kable("html", escape = FALSE, caption = "Chart length by specialty and language") |>
  style_table()
```

```{r chart-length-plot, fig.cap="Chart token counts by specialty and language."}
chart_lengths |>
  ggplot(aes(x = language, y = chartTokenCount, fill = specialty)) +
  geom_boxplot(alpha = 0.8, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.5) +
  labs(x = "Language", y = "Approximate chart token count", fill = "Specialty") +
  theme_minimal()
```

# PDQI+ Quality

```{r quality-summary}
reviews |>
  group_by(generator, language) |>
  summarise(
    Reviews = n(),
    `Overall score` = mean_ci(overall),
    `PDQI-9 mean` = mean_ci(pdqi9_mean),
    Safety = mean_ci(safety),
    Bias = mean_ci(bias),
    `Hallucination severity` = mean_ci(hallucinationSeverity),
    .groups = "drop"
  ) |>
  knitr::kable("html", escape = FALSE, caption = "Quality outcomes by generator and language") |>
  style_table()
```

```{r pdqi-domain-summary}
reviews |>
  select(generator, language, all_of(pdqi_cols)) |>
  pivot_longer(cols = all_of(pdqi_cols), names_to = "domain", values_to = "score") |>
  mutate(domain = str_remove(domain, "^pdqi_"), domain = recode(domain, !!!pdqi_labels)) |>
  group_by(generator, language, domain) |>
  summarise(Score = mean(score, na.rm = TRUE), .groups = "drop") |>
  mutate(Score = txt_round(Score, 2)) |>
  pivot_wider(names_from = language, values_from = Score) |>
  arrange(generator, domain) |>
  knitr::kable("html", escape = FALSE, caption = "Mean PDQI-9 domain scores") |>
  style_table()
```

```{r overall-plot, fig.cap="Overall score distribution by generator and language."}
reviews |>
  filter(!is.na(overall)) |>
  ggplot(aes(x = factor(overall), fill = generator)) +
  geom_bar(position = "dodge", color = "black", alpha = 0.8) +
  facet_wrap(~language) +
  labs(x = "Overall score", y = "Review count", fill = "Generator") +
  theme_minimal()
```

```{r pdqi-chart-length-plot, fig.cap="PDQI-9 mean score by chart token count."}
reviews |>
  filter(!is.na(pdqi9_mean), !is.na(chartTokenCount)) |>
  ggplot(aes(x = chartTokenCount, y = pdqi9_mean, color = generator)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Approximate chart token count", y = "PDQI-9 mean", color = "Generator") +
  theme_minimal()
```

```{r prompt-summary}
reviews |>
  filter(generator == "LLM") |>
  group_by(promptType, summaryGeneratedBy) |>
  summarise(
    Reviews = n(),
    `Overall score` = mean_ci(overall),
    `PDQI-9 mean` = mean_ci(pdqi9_mean),
    Safety = mean_ci(safety),
    Bias = mean_ci(bias),
    `Hallucination severity` = mean_ci(hallucinationSeverity),
    .groups = "drop"
  ) |>
  arrange(desc(Reviews), summaryGeneratedBy) |>
  knitr::kable("html", escape = FALSE, caption = "LLM prompt/model subgroup quality") |>
  style_table()
```

# Mixed Models

Mixed models include generator type and scaled chart token count as fixed effects, with reviewer as a random intercept. Models are omitted when the selected campaign has too few observations or insufficient variation.

```{r mixed-models}
quality_outcomes <- c(
  overall = "overall score",
  pdqi9_mean = "PDQI-9 mean",
  safety = "safety",
  bias = "bias",
  hallucinationSeverity = "hallucination severity"
)

models <- imap(quality_outcomes, \(label, outcome) {
  if (!requireNamespace("lme4", quietly = TRUE)) {
    return(glue("Mixed model for {label} was not fitted because lme4 is not installed."))
  }

  model <- fit_quality_model(reviews, outcome)
  if (is.null(model)) {
    return(glue("Mixed model for {label} was not fitted because the campaign has too few observations or insufficient variation."))
  }

  output_mixed_model(model, label)
})

for (model_output in models) {
  print(model_output)
}
```

# Comments

```{r low-score-comments}
comment_cols <- c(
  "overallComment",
  "safety__comment",
  "bias__comment",
  "hallucinationSeverity__comment",
  paste0("pdqi_", pdqi_domains, "__comment")
)
available_comment_cols <- intersect(comment_cols, names(reviews))

reviews |>
  select(generator, language, specialty, caseName, summaryGeneratedBy, reviewerName,
         overall, pdqi9_mean, all_of(available_comment_cols)) |>
  pivot_longer(cols = all_of(available_comment_cols),
               names_to = "commentType",
               values_to = "comment") |>
  filter(!is.na(comment), str_trim(comment) != "") |>
  mutate(commentType = str_replace_all(commentType, "_", " ")) |>
  arrange(generator, language, specialty, caseName) |>
  knitr::kable("html", escape = FALSE, caption = "Reviewer comments") |>
  style_table()
```

# Session

```{r session-info}
sessionInfo()
```
